---
title: "scRNAseq Simulations"
author: "Shirley Mathur"
date:   "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages, eval = TRUE}
#load packages
library(tidyverse)
library(Matrix)
library(MASS)
library(Rcpp)
library(RcppArmadillo)
library(nloptr)
library(numDeriv)
library(PLNmodels)
library(patchwork)
```

```{r load-functions, eval = TRUE}
#import functions
source("scrnaseq_project_functions.R")
sourceCpp('scrnaseq_project_cpp_functions.cpp')
```

## Simulation Settings

```{r sim-setup, eval = TRUE}
#set fixed parameter values
set.seed(12)

#set number of samples
n <- 100
#set number of timepoints per sample
m <- 5
#set number of categories
J <- 10

#set VAR(1) noise covariance(Sigma), VAR coefficient matrix (A)
#Sigma <- matrix(c(0.5, 0.3, 0.3, 0.5), nrow = J, ncol = J) 
Sigma <- diag(0.5, nrow = J, ncol = J)
#A <- matrix(c(0.5, rep(0,J^2 - 1)), nrow = J)
#A <- matrix(0.1*1:J^2, nrow = J)
#A = matrix(c(0.1,0.5,0.3,0.7),2,2)
A <- matrix(sparseMatrix(i = 1:J, j = sample(1:J, J), x = 0.5, dims = c(J,J)), J, J)
Sigma_Z <- matrix(solve(diag(1, J*J, J*J) - kronecker(A, A)) %*% c(Sigma),
       nrow = J, ncol = J)


#Set mean offset in parameters of Poisson distribution for observed values
#mu <- matrix(runif(J), nrow = J, ncol = 1)
#mu <- matrix(rep(1.5,J), nrow = J)
mu <- matrix(rep(0.5, J), nrow = J)

#setup for simulation
nsim <- 5
elbo2 <- TRUE #determine if ELBO based on likelihood conditional on first timepoint or not will be used (will be used if TRUE)
penalty <- TRUE #determine whether to use penalized estimators or not
if (elbo2) {full_params <- c("mu", "Sigma", "A")} else { full_params <- c("mu", "Sigma_Z", "A") }
skip_params <- NA

sim_sample_list <- vector(mode = "list")
sim_mom_list <- vector(mode = "list")

sim_mu_results <- array(NA, dim = c(2,J,nsim),
                     dimnames = list("est_method" = c("mom", "vi"),
                                  "est" = paste0("mu", 1:J),
                                  "iter" = 1:nsim))

sim_Sigma_Z_results <- array(NA, dim = c(2,J^2,nsim),
                        dimnames = list("est_method" = c("mom", "vi"),
                                        "est" = paste0("SZ_", paste0(rep(1:J,J), rep(1:J, each = J))),
                                        "iter" = 1:nsim))

sim_Sigma_results <- array(NA, dim = c(2,J^2,nsim),
                        dimnames = list("est_method" = c("mom", "vi"),
                                        "est" = paste0("Sigma_", paste0(rep(1:J,J), rep(1:J, each = J))),
                                        "iter" = 1:nsim))

sim_A_results <- array(1:(2*J^2*nsim), dim = c(2,J, J,nsim),
                             dimnames = list("est_method" = c("mom", "vi"),
                                             #"est" = paste0("A_", paste0(rep(1:J,J), rep(1:J, each = J))),
                                             "row" = 1:J,
                                             "column" = 1:J,
                                             "iter" = 1:nsim))

A_max_eig_mod <- rep(NA, nsim)

sim_Sigma_list <- vector(mode = "list")
sim_Sigma_pd <- rep(NA, nsim)

sim_M_list <- vector(mode = "list")
sim_S_list <- vector(mode = "list")

#for saving sim settings
sim_settings <- list("n" = n,
                     "m" = m,
                     "J" = J,
                     "nsim" = nsim,
                     "mu" = mu,
                     "Sigma_Z" = Sigma_Z,
                     "A" = A,
                     "opt_params" = setdiff(full_params, skip_params))
```

-   Sample size (n): `r n`
-   Number of timepoints (m): `r m`
-   Dimension of outcome (J): `r J`
-   Number of simulation runs: `r nsim`
-   Parameters optimized over: `r setdiff(full_params, skip_params)`
-   ELBO Utilized: `r if_else(elbo2, "Conditional on Z_1", "Original")`
-   Penalty:
    `r if_else(penalty, "penalized estimators", "non-penalized estimators")`

## Simulation Results

```{r run-sims}
for (i in 1:nsim) {
  #print(i)
  #print iteration for every 50 sims
  # if (i %% 50 == 0) {
  #   print(paste0("Simulation iteration: ", i))
  # }
  
  #generate sample data
  temp_sample <- sim_data(n, m, J, Sigma, A, mu)
  sample_list[[i]] <- temp_sample
  
  #COMPUTE MoM ESTIMATORS and store result
  lambda <- seq(1e-2,1, length.out = 100)
  mom_estimates <- vector(mode = "list")
  for (l in 1:length(lambda)) {
    mom_estimates[[l]] <- mom_estimator(temp_sample$Y, penalty = penalty, lambda = lambda[l])
  }
  sim_mom_list[[i]] <- list("lambda" = lambda, "estimates" = mom_estimates)
  # sim_mu_results["mom", ,paste0(i)] <- mu_hat <- mom_estimates$mu_hat
  # Sigma_Z_hat_vec <- c(mom_estimates$Sigma_Z_hat)
  # sim_Sigma_Z_results["mom", ,paste0(i)] <- c(mom_estimates$Sigma_Z_hat)
  # Sigma_hat_vec <- c(mom_estimates$Sigma_hat)
  # sim_Sigma_results["mom", , paste0(i)] <- c(mom_estimates$Sigma_hat)
  # A_hat_vec <- c(mom_estimates$A_hat)
  # sim_A_results["mom", , ,paste0(i)] <- mom_estimates$A_hat

  #COMPUTE VI ESTIMATES and store result
  if (!elbo2) {
    vi_est <- vi_estimator(temp_sample,
                         init_mu = mu_hat,
                         init_Sigma_Z = Sigma_Z_hat_vec,
                         init_A = A_hat_vec,
                         init_M = rep(0, m*J*n),
                         init_S = rep(1, m*J*n),
                         optim_method = "nloptr",
                         max.iter = 100,
                         skip_coords = skip_params)

    sim_Sigma_Z_results["vi", ,paste0(i)] <- c(vi_est$Sigma_Z)

  } else {
    #compute warm start values with non-penalized MoM estimates
    mom_estimates_nopen <-  mom_estimator(temp_sample$Y, penalty = FALSE)
    A_hat_vec <- c(mom_estimates_nopen$A_hat)
    mu_hat <- mom_estimates_nopen$mu_hat
    Sigma_hat_vec <- c(mom_estimates_nopen$Sigma_hat)

    #not getting fixed M and S first
    vi_estimates <- vector(mode = "list")
    lambda <- seq(1e-3, 1, length.out = 1000)
    for (l in 1:length(lambda)) {
      print(l)
      vi_estimates[[l]] <- vi_estimator2(temp_sample,
                        init_mu = mu_hat,
                        init_Sigma = Sigma_hat_vec,
                        init_A = A_hat_vec,
                        init_M = rep(0, m*J*n),
                        init_S = rep(1, m*J*n),
                        optim_method = "nloptr",
                        max.iter = 100,
                        penalty = penalty,
                        lambda = lambda[l],
                        verbose = FALSE,
                        skip_coords = skip_params)
    }

    vi_est <- vi_estimator2(temp_sample,
                        init_mu = mu_hat,
                        init_Sigma = Sigma_hat_vec,
                        init_A = A_hat_vec,
                        init_M = rep(0, m*J*n),
                        init_S = rep(1, m*J*n),
                        optim_method = "nloptr",
                        max.iter = 100,
                        penalty = penalty,
                        lambda = 1e-2,
                        verbose = FALSE,
                        skip_coords = skip_params)


    # #getting M and S fixed first
    # vi_est_init <- vi_estimator2(temp_sample,
    #                      init_mu = mu_hat,
    #                      init_Sigma = Sigma_hat_vec,
    #                      init_A = A_hat_vec,
    #                      init_M = rep(0, m*J*n),
    #                      init_S = rep(1, m*J*n),
    #                      optim_method = "nloptr",
    #                      max.iter = 100,
    #                      skip_coords = full_params)
    #
    # M_fix <- c(vi_est_init$M)
    # S_fix <- c(vi_est_init$S)
    #
    # vi_est <- vi_estimator2(temp_sample,
    #                      init_mu = mu_hat,
    #                      init_Sigma = Sigma_hat_vec,
    #                      init_A = A_hat_vec,
    #                      init_M = M_fix,
    #                      init_S = S_fix,
    #                      optim_method = "nloptr",
    #                      max.iter = 100,
    #                      penalty = TRUE,
    #                      lambda = 1e-3,
    #                      skip_coords = c("M", "S", skip_params))

    sim_Sigma_results["vi", ,paste0(i)] <- c(vi_est$Sigma)
    sim_M_list[[i]] <- vi_est$M
    sim_S_list[[i]] <- vi_est$S
  }


  sim_mu_results["vi", ,paste0(i)] <- vi_est$mu
  sim_A_results["vi", , ,paste0(i)] <- vi_est$A
  A_max_eig_mod[i] <- max(Mod(eigen(vi_est$A)$values))
  
}
```

```{r run-sims-chiquet}
for (i in 1:nsim) {
  #generate sample data
  temp_sample <- sim_data(n, m, J, Sigma, A, mu)
  sample_list[[i]] <- temp_sample
  
  #prepare it so you can fit PLN model to it
  temp_list <- list("counts" = cbind(c(temp_sample$Y[,1,]), c(temp_sample$Y[,2,])),
                    "covariates" = matrix(1, nrow = m*n, ncol = 1))
  
  rownames(temp_list$counts) <- rownames(temp_list$covariates) <- c(1:(m*n))
  colnames(temp_list$counts) <- c("Y1", "Y2")
  
  temp_prep_data <- prepare_data(temp_list$counts, temp_list$covariates, offset = "none")

  
  #COMPUTE MoM ESTIMATORS and store result
  mom_estimates <- mom_estimator(temp_sample$Y)
  sim_mu_results["mom", ,paste0(i)] <- mu_hat <- mom_estimates$mu_hat
  sim_Sigma_Z_results["mom", ,paste0(i)] <- Sigma_Z_hat_vec <- c(mom_estimates$Sigma_Z_hat)
  sim_Sigma_results["mom", , paste0(i)] <- Sigma_hat_vec <- c(mom_estimates$Sigma_hat)
  sim_A_results["mom", ,paste0(i)] <- A_hat_vec <- c(mom_estimates$A_hat)
  
  #COMPUTE VI ESTIMATES and store result
  if (!elbo2) {
    vi_est <- vi_estimator(temp_sample, 
                         init_mu = mu_hat, 
                         init_Sigma_Z = Sigma_Z_hat_vec,
                         init_A = A_hat_vec,
                         init_M = rep(0, m*J*n),
                         init_S = rep(1, m*J*n),
                         optim_method = "nloptr",
                         max.iter = 100,
                         skip_coords = skip_params)
    
    sim_Sigma_Z_results["vi", ,paste0(i)] <- c(vi_est$Sigma_Z)
    
  } else {
    # vi_est_init <- vi_estimator2(temp_sample,
    #                      init_mu = mu_hat,
    #                      init_Sigma = Sigma_hat_vec,
    #                      init_A = A_hat_vec,
    #                      init_M = rep(0, m*J*n),
    #                      init_S = rep(1, m*J*n),
    #                      optim_method = "nloptr",
    #                      max.iter = 100,
    #                      skip_coords = full_params)
    # 
    # M_fix <- c(vi_est_init$M)
    #S_fix <- c(vi_est_init$S)

    vi_est <- vi_estimator2(temp_sample, 
                         init_mu = mu_hat, 
                         init_Sigma = Sigma_hat_vec,
                         init_A = A_hat_vec,
                         init_M = rep(0, m*J*n),
                         init_S = rep(1, m*J*n),
                         optim_method = "nloptr",
                         max.iter = 100,
                         tol = 1e-8,
                         skip_coords = skip_params)
    
    sim_Sigma_results["vi", ,paste0(i)] <- c(vi_est$Sigma)
    sim_M_list[[i]] <- vi_est$M
    sim_S_list[[i]] <- vi_est$S
  }
  
  
  sim_mu_results["vi", ,paste0(i)] <- vi_est$mu
  sim_A_results["vi", ,paste0(i)] <- c(vi_est$A)
  A_max_eig_mod[i] <- max(Mod(eigen(vi_est$A)$values))
  
  #COMPUTE PLN ESTIMATES AND STORE RESULT
  pln_fit <- PLN(Abundance ~ 1, temp_prep_data, control = PLN_param(trace = 1))
  sim_mu_results["chiquet", ,paste0(i)] <- c(pln_fit$model_par$Theta)
  sim_Sigma_results["chiquet",,paste0(i)] <- c(pln_fit$model_par$Sigma)
  
}
```

```{r report-results}
#save results
save(sim_settings, sim_mu_results, sim_A_results, sim_Sigma_results, sim_Sigma_Z_results, A_max_eig_mod, sim_M_list, sim_S_list, sample_list, file = "all_params_J2_sparseA_penalized_sim.Rdata")

#tabulate results
sim_res_df <- as.data.frame.table(sim_mu_results, responseName = "value")
sim_SZ_res_df <- as.data.frame.table(sim_Sigma_Z_results, responseName = "value")
sim_A_res_df <- as.data.frame.table(sim_A_results, responseName = "value")
sim_Sigma_res_df <- as.data.frame.table(sim_Sigma_results, responseName = "value")

#data frame of mu results
mu_summary <- sim_res_df %>% group_by(est_method, est) %>%
  summarise(mean_val = mean(value),
            median_val = median(value),
            sd = sd(value)) %>%
  mutate(true_val = case_when(est == "mu1" ~ mu[1],
                              est == "mu2" ~ mu[2],
                              est == "mu3" ~ mu[3],
                              est == "mu4" ~ mu[4])) %>%
  relocate(true_val, .before = mean_val)

#data frame of Sigma_Z results
sigma_Z_summary <- sim_SZ_res_df %>% group_by(est_method, est) %>%
  summarise(mean_val = mean(value),
            median_val = median(value),
            sd = sd(value)) %>%
  mutate(true_val = case_when(est == "SZ_11" ~ Sigma_Z[1,1],
                              est == "SZ_12" ~ Sigma_Z[1,2],
                              est == "SZ_21" ~ Sigma_Z[2,1],
                              est == "SZ_22" ~ Sigma_Z[2,2])) %>%
  relocate(true_val, .before = mean_val)

#data frame of Sigma results
sigma_summary <- sim_Sigma_res_df %>% group_by(est_method, est) %>%
  summarise(mean_val = mean(value),
            median_val = median(value),
            sd = sd(value)) %>%
  mutate(true_val = case_when(est == "Sigma_11" ~ Sigma[1,1],
                              est == "Sigma_12" ~ Sigma[1,2],
                              est == "Sigma_21" ~ Sigma[2,1],
                              est == "Sigma_22" ~ Sigma[2,2])) %>%
  relocate(true_val, .before = mean_val)

#data frame of A results
A_summary <- sim_A_res_df %>% 
  mutate(index = paste0(row, column)) %>% 
  group_by(est_method, index) %>%
  summarise(mean_val = mean(value),
            median_val = median(value),
            sd = sd(value),
            prop_zero = mean(value == 0)) 
  # mutate(true_val = case_when(est == "A_11" ~ A[1,1],
  #                          est == "A_12" ~ A[1,2],
  #                          est == "A_21" ~ A[2,1],
  #                          est == "A_22" ~ A[2,2]),
  #        prop_nonzero = 1 - prop_zero) %>%
  # relocate(true_val, .before = mean_val)

all_summary <- data.frame()

if (!("mu" %in% skip_params)) {
  all_summary <- rbind(all_summary, mu_summary)
}

if (!("Sigma_Z" %in% skip_params) & !elbo2) {
  all_summary <- rbind(all_summary, sigma_Z_summary)
}

if (!("A" %in% skip_params)) {
  all_summary <- rbind(all_summary, A_summary)
}

if (!("Sigma" %in% skip_params) & elbo2) {
  all_summary <- rbind(all_summary, sigma_summary)
}


all_summary %>% kableExtra::kable(digits = 3, caption = "Summary of Results")

# #plot of mu errors
# png("mu_errors.png", width = 6, height = 6, units = "in", res = 300)
if (!("mu" %in% skip_params)) {
  sim_res_df %>% 
    mutate(error = case_when(est == "mu1" ~ value - mu[1],
                             est == "mu2" ~ value - mu[2],
                             est == "mu3" ~ value - mu[3],
                             est == "mu4" ~ value - mu[4]))%>%
    ggplot(mapping = aes(x = error)) +
    geom_boxplot() +
    facet_grid(est_method ~ est) +
    labs(title = "Summary of Errors Across Simulations for mu",
         x = "Error (Est - True)") +
    theme_bw() +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
  # dev.off()
}  


# #plot of Sigma_Z errors
if (!("Sigma_Z" %in% skip_params) & !elbo2) {
  # png("Sigma_Z_errors.png", width = 6, height = 6, units = "in", res = 300)
  sim_SZ_res_df %>%
    mutate(error = case_when(est == "SZ_11" ~ value - Sigma_Z[1,1],
                             est == "SZ_12" ~ value - Sigma_Z[1,2],
                             est == "SZ_21" ~ value - Sigma_Z[2,1],
                             est == "SZ_22" ~ value - Sigma_Z[2,2])) %>%
    ggplot(mapping = aes(x = error)) +
    geom_boxplot() +
    facet_grid(est_method ~ est, scales = "free") +
    labs(title = "Summary of Errors Across Simulations for Sigma_Z",
         x = "Error (Est - True)") +
    theme_bw() +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
  
  # dev.off()
}

#plot of A errors
if (!("A" %in% skip_params)) {
  sim_A_res_df %>%
  mutate(error = case_when(est == "A_11" ~ value - A[1,1],
                           est == "A_12" ~ value - A[1,2],
                           est == "A_21" ~ value - A[2,1],
                           est == "A_22" ~ value - A[2,2])) %>%
  ggplot(mapping = aes(x = error)) +
  geom_boxplot() +
  facet_grid(est_method ~ est, scales = "free") +
  labs(title = "Summary of Errors Across Simulations for A",
       x = "Error (Est - True)") +
  theme_bw() +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
}

#plot of Sigma errors
if (!("Sigma" %in% skip_params) & elbo2) {
  sim_Sigma_res_df %>%
  mutate(error = case_when(est == "Sigma_11" ~ value - Sigma[1,1],
                           est == "Sigma_12" ~ value - Sigma[1,2],
                           est == "Sigma_21" ~ value - Sigma[2,1],
                           est == "Sigma_22" ~ value - Sigma[2,2])) %>%
  ggplot(mapping = aes(x = error)) +
  geom_boxplot() +
  facet_grid(est_method ~ est, scales = "free") +
  labs(title = "Summary of Errors Across Simulations for Sigma",
       x = "Error (Est - True)") +
  theme_bw() +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
}
```

```{r mom-model-selection-sims}
nsim <- 5
for (i in 1:nsim) {
  print(i)
  #generate sample data
  temp_sample <- sim_data(n, m, J, Sigma, A, mu)
  sim_sample_list[[i]] <- temp_sample
  
  #COMPUTE MoM ESTIMATORS and store result
  lambda <- seq(1e-3,1, length.out = 1000)
  mom_estimates <- vector(mode = "list")
  
  for (l in 1:length(lambda)) {
    mom_estimates[[l]] <- mom_estimate_temp <- mom_estimator(temp_sample$Y, penalty = penalty, lambda = lambda[l])
  }
  sim_mom_list[[i]] <- list("lambda" = lambda, "estimates" = mom_estimates)
}

```


```{r load-lambda-selection-results, eval = TRUE}
#load appropriate results and extract relevant information
#load("vi_ebic_penalty_mom_model_selection.Rdata")
#load("exploring_penalty_mom_model_selection.Rdata")
load("Simulation Results/mom_penalized_model_selection_100_sims.Rdata")
A <- sim_settings$A
nsim <- length(sim_mom_list)
n <- sim_settings$n
J <- sim_settings$J
m <- sim_settings$m
lambdas <- sim_mom_list[[1]]$lambda
n_lambda <- length(lambdas)
```

```{r lambda-selection-results, eval = TRUE}
#get true support of A
A_true_support <- which(A != 0)

#initialize data frame to store edges recovered, fpr, and tpr for each lambda and each sim
pdata_list <- vector(mode = "list")

for (i in 1:nsim) {
  est_support <- lapply(sim_mom_list[[i]]$estimates, function(x) {which(x$A != 0)})
  tpr <- sapply(est_support, function (x) {length(intersect(x, A_true_support))/length(A_true_support)})
  fpr <- sapply(est_support, function (x) {ifelse(length(x) > 0, length(setdiff(x, A_true_support))/(J^2 - length(A_true_support)), 0)})
  num_edges <- sapply(est_support, function (x) {length(x)})
  #print(which(tpr == 1))
  temp_pdata <- data.frame(lambda = sim_mom_list[[i]]$lambda,
                      fpr = fpr,
                      num_edges = num_edges,
                      tpr = tpr,
                      sim = i)
  pdata_list[[i]] <- temp_pdata
  
  #make plots of first 5 sims only
  if (i <= 5) {
    temp_plot <- temp_pdata %>% ggplot(mapping = aes(x = num_edges, y = tpr)) +
    geom_point(alpha = 0.5) +
    labs(title = paste0("TPR in Sim ", i), 
         x = "Number Edges Recovered",
         y = "True Positive Rate") +
    ylim(0, 1.2) +
    xlim(0, 20) +
    theme_bw()
  
  assign(x = paste0("p",i), value = temp_plot)
  }
  
  #  plot(sim_mom_list[[i]]$lambda, tpr, xlab = "lambda", ylab = "TPR", main = paste0("TPR in Sim ", i))
}

edge_recovery_metrics_full <- do.call(rbind, pdata_list)


#add evaluations of various evaluation criteria to dataframe with fpr and tpr to try and understand what might work
# for model selection in this scenario

#list out ess and criteria types you want to assess
ess_types <- c("ss", "ess", "ess2", "ess3", "ess4")
criteria_types <- c("mse", "bic", "ebic", "bic2", "ebic2")

#set up array to store info on: simulation run, lambda, ess computation approach, and model selection criteria type
model_selection_results <- array(NA, dim = c(nsim, n_lambda,length(ess_types), length(criteria_types)),
                                 dimnames = list("sim" = 1:nsim,
                                                 "lambda" = lambdas,
                                                 "ess_type" = ess_types,
                                                 "criteria" = criteria_types))

ess_values <- array(NA, dim = c(nsim, length(ess_types)),
                    dimnames = list("sim" = 1:nsim,
                                    "ess_type" = ess_types))

for (i in 1:nsim) {
  
  #compute terms necessary to estimate autocorrelation for lags of k=1, 2, ..., T-1
  Y_samp <- sim_sample_list[[i]]$Y
  Y_means <- apply(Y_samp, 2, mean)
  Y_vars <- apply(Y_samp, 2, function (x) {var(c(x))})
  Y_samp_centered <- sweep(Y_samp, 2, Y_means)
  rho_1_est <- apply(Y_samp_centered[1:(m-1),,]*Y_samp_centered[2:m,,], 2, mean)/Y_vars
  rho_k_full <- apply(matrix(1:(m-1), 1), 2, function (x) {(1-x/m)*rho_1_est^x})
  tau_full <- apply(rho_k_full, 1, function (x) {1 + 2*sum(x)})
  
  #COMPUTE MODEL SELECTION METRICS FOR EACH DIFFERENT ESS COMPUTATION APPROACH
  for (ess_type in ess_types) {
    #compute ess accordingly and store in ess_values array
    ess_values[i, ess_type] <- N_eff <- case_when(ess_type == "ss" ~ n*(m-1),
                       ess_type == "ess" ~ n*(m-1)/(1 + 2*max(rho_1_est)),
                       ess_type == "ess2" ~ n*(m-1)/(1 + 2*mean(rho_1_est)),
                       ess_type == "ess3" ~ n*(m-1)/(max(tau_full)),
                       ess_type == "ess4" ~ n*(m-1)/(mean(tau_full)))
    
    #compute model selection criteria for current ess type
    for (crit in criteria_types) {
      if (crit == "mse") {
        model_selection_results[i, , ess_type, crit] <- sapply(sim_mom_list[[i]]$estimates, function(x) {norm(x$A_hat - x$P %*% solve(x$Sigma_Z_hat), type = "F")})
      }
      if (crit == "bic") {
        model_selection_results[i, , ess_type, crit] <- sapply(sim_mom_list[[i]]$estimates, function(x) {N_eff*norm(x$A_hat - x$P %*% solve(x$Sigma_Z_hat), type = "F") + log(N_eff)*sum(x$A_hat != 0)})
      }
      
      if (crit == "ebic") {
        model_selection_results[i, , ess_type, crit] <- sapply(sim_mom_list[[i]]$estimates, function(x) {N_eff*norm(x$A_hat - x$P %*% solve(x$Sigma_Z_hat), type = "F") + log(N_eff)*sum(x$A_hat != 0) + log(choose(J^2, sum(x$A_hat != 0)))})
      }
      
      if (crit == "bic2") {
        model_selection_results[i, , ess_type, crit] <- sapply(sim_mom_list[[i]]$estimates, function(x) {N_eff*norm(x$A_hat %*% x$Sigma_Z_hat - x$P, type = "F") + log(N_eff)*sum(x$A_hat != 0)})
      }
      
      if (crit == "ebic2") {
        model_selection_results[i, , ess_type, crit] <- sapply(sim_mom_list[[i]]$estimates, function(x) {N_eff*norm(x$A_hat %*% x$Sigma_Z_hat - x$P, type = "F") + log(N_eff)*sum(x$A_hat != 0) + log(choose(J^2, sum(x$A_hat != 0)))})
      }
      
    }
  }
}

#transfer array results into a data frame
model_selection_results_df <- as.data.frame.table(model_selection_results, responseName = "value")
```

Below are results for the true positive rate and the number of edges
recovered for the first 5 sims across a grid of `r n_lambda` lambda
values with increments of `r lambdas[2] - lambdas[1]` from
`r lambdas[1]` to `r sim_mom_list[[1]]$lambda[n_lambda]`.

```{r results-from-specific-sims, eval = TRUE}
#plot fpr and edges recovered plots for first five sims
(p1 + p2 + p3)/(p4 + p5)

# # plot out ROC curves along with colored gradient for various selection criteria for first 5 sims
# pdata_full %>% filter(sim <= 5) %>%
#   ggplot(mapping = aes(x = fpr, y = tpr, color = mse)) +
#   geom_point() +
#   labs(title = "ROC Curves for Simulation and MSE",
#        xlab = "FPR",
#        ylab = "TPR",
#        color = "MSE") +
#   facet_wrap(~ sim) +
#   theme_bw()
# 
# # plot out fpr and ebic
# pdata_full %>% filter(sim <= 5) %>%
#   ggplot(mapping = aes(x = fpr, y = tpr, color = bic_ess)) +
#   geom_point() +
#   labs(title = "ROC Curves for Simulation and BIC with ESS",
#        xlab = "FPR",
#        ylab = "TPR",
#        color = "EBIC") +
#   facet_wrap(~ sim) +
#   theme_bw()
# 
# # plot out fpr and ebic
# pdata_full %>% filter(sim <= 5) %>%
#   ggplot(mapping = aes(x = fpr, y = tpr, color = ebic_ess)) +
#   geom_point() +
#   labs(title = "ROC Curves for Simulation and EBIC",
#        xlab = "FPR",
#        ylab = "TPR",
#        color = "EBIC") +
#   facet_wrap(~ sim) +
#   theme_bw()

# pdata_full %>%
#   ggplot(mapping = aes(x = fpr, y = tpr, color = vi_ebic)) +
#   geom_point() +
#   labs(title = "ROC Curves for Simulation and VI EBIC",
#        xlab = "FPR",
#        ylab = "TPR",
#        color = "VI_EBIC") +
#   facet_wrap(~ sim) +
#   theme_bw()
```

Some notation for clarity:

-   There $N$ samples, $T$ time-points, and $J$ categories

-   $\bar{Y}_j := \frac{1}{NT}\sum\limits_{i=1}^N\sum\limits_{t=1}^{T} Y_{ij}^{(t)}$

-   $\hat{\sigma}_j^2 :=  \frac{1}{NT - 1}\sum\limits_{i=1}^N\sum\limits_{t=1}^{T} \left(Y_{ij}^{(t)} - \bar{Y}_j\right)^2$

-   $\hat{\rho}_j :=  \frac{1}{\hat{\sigma}_j^2}\frac{1}{N(T-1)}\sum\limits_{i=1}^N\sum\limits_{t=1}^{T-1}\left(Y^{(t)}_{ij} - \bar{Y}_j \right)\left(Y^{(t+1)}_{tj} - \bar{Y}_j \right)$

-   ${\hat{\rho}}_{\max} := \underset{j \in 1:J}{\max}\hat{\rho}_j$

-   ${\hat{\rho}}_{\text{mean}} := \frac{1}{J}\sum\limits_{j=1}^J\hat{\rho}_j$

-   $\tau_j := 1 + 2\sum\limits_{k=1}{T-1} \left(1 - \frac{k}{T} \right)\hat{\rho}_j^{k}$

-   $\tau_{\max} := \underset{1 \leq j \leq J}{\max} \tau_j$

-   $\tau_{\text{mean}} := \frac{1}{J}\sum\limits_{j=1}^J \tau_j$

In the below, the selection criteria are computed as follows:

-   `bic` and `ebic` utilize $A - P\Sigma_Z^{-1}$ as the moment matching
    criterion

-   `bic2` and `ebic2` utilize $A\Sigma_Z - P$ as the moment matching
    criterion

-   `ess` specifies that the effective sample size is computed as
    $\text{ESS} = N(T-1)\left(1 - 2{\hat{\rho}}_{\max} \right)^{-1}$

-   `ess2` specifies that the effective sample size is computed as
    $\text{ESS} = N(T-1)\left(1 - 2{\hat{\rho}}_{\text{mean}} \right)^{-1}$

-   `ess3` specifies that the effective sample size is computed as
    $\text{ESS} = \frac{N(T-1)}{\tau_{\max}}$

-   `ess4` specifies that the effective sample size is computed as
    $\text{ESS} = \frac{N(T-1)}{\tau_{\text{mean}}}$

```{r selection-results-summary, eval = TRUE}
#selected models according to different criteria across all simulations in one dataframe
selection_sims_summary <- model_selection_results_df %>%
  mutate(sim = as.numeric(sim)) %>%
  group_by(sim, ess_type, criteria) %>%
  summarise(min_val = min(value),
            which_min = which.min(value)) %>%
  mutate(lambda = lambdas[which_min]) %>%
  dplyr::select(!c(which_min))
  
selection_sims_summary_full <- left_join(selection_sims_summary, edge_recovery_metrics_full) %>%
  rename(num_edges_criteria = num_edges, tpr_criteria = tpr)


#in summary frame, from full results list, add in empirically which was the lambda with the smalles number of edges that gave a tpr of 1
edge_recovery_summary <- edge_recovery_metrics_full %>%
  filter(tpr == 1) %>%
  group_by(sim) %>%
  summarise(minimal_edges = min(num_edges),
            which_min = which.min(num_edges)) %>%
  mutate(lambda = lambdas[which_min]) %>%
  dplyr::select(!c(which_min))

# see which, if any, criteria actually select the lambda that empirically had the smallest number of edges and a tpr of 1 across the simulations
criteria_comparison_results <- left_join(selection_sims_summary_full, edge_recovery_summary, by = c("sim"), suffix = c("_criteria", "_empirical"))

criteria_comparison_results_summary <- criteria_comparison_results %>% 
  ungroup() %>% 
  group_by(criteria) %>% 
  summarise(median_lambda_diff = median(lambda_empirical - lambda_criteria),
            median_edges_diff = median(minimal_edges - num_edges_criteria))


#visualize edges recovered from all selection methods
selection_sims_summary_full %>% ggplot(mapping = aes(x = num_edges_criteria)) +
  geom_histogram() +
  facet_grid(criteria ~ ess_type) +
  labs(title = "Histograms of Edges Recovered with Various Selection Criteria",
       x = "Edges Recovered",
       caption = "Histograms of the edges recovered by the model selected by the various selection criteria across 100 simulations.") +
  theme_bw()

selection_sims_summary_full %>% ggplot(mapping = aes(x = num_edges_criteria, y = tpr_criteria)) +
  geom_point() +
  facet_grid(criteria ~ ess_type) +
  labs(title = "Edges Recovered and TPR across Sims with Various Selection Criteria",
       x = "Edges Recovered",
       y = "TPR",
       caption = "TPR and edges recovered by the model selected by the various selection criteria across 100 simulations.") +
  theme_bw()

criteria_comparison_results %>% mutate(minimal_edges_diff = num_edges_criteria - minimal_edges) %>%
  ggplot(mapping = aes(x = minimal_edges_diff)) +
  geom_boxplot() +
  facet_grid(criteria ~ ess_type) +
  labs(title = "Difference in empirical minimial edges for tpr of 1 and num edges with selected model",
       x = "edges difference (selected model - empirical best model)",
       caption = "Difference in selected model edges and empirical minimial edges for TPR of 1 by the various selection criteria across 100 simulations.") +
  theme_bw()
  
```
